---
title: "brownfat"
output: html_document
---

```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(pheatmap)
library(car)
```

Read the file:

```{r}
# the excel sheet represents NA values by "NA"
brownfat <- read_excel("./description/BrownFat.xls", na = "NA")
nrow(brownfat)
```

Check if there's any NA values:

```{r}
x <- colSums(is.na(brownfat)) # 
x[x > 0]
```

3 variables has NA values detected: Cancer_Status, Cancer_Type, and TSH. Data cleaning needed.

Notice that $4425$ out of $n=4842$ observations had no data on their `TSH`, as this ratio is too large in order for us to confidently assign the most-probable values, we will choose to exclude the variable `TSH` altogether.

```{r}
brownfat_cleaned_data <- brownfat %>% select(-c("TSH", "Id"))
```

There are observations where both the `Cancer_Status` and `Cancer_Type` is set to NA, as we don't have domain knowledge to assign appropriate values nor is it the main focus of this research, we will simply exclude these observations.

```{r}
# determine how many obs. display such behaviour
x <- sum(is.na(brownfat$Cancer_Status) & is.na(brownfat$Cancer_Type))
paste("No. of observations w/ Cancer_Status == Cancer_Type == NA is:", x)

# exclude these obs.
brownfat_cleaned_data <- brownfat_cleaned_data %>% 
  filter(!(is.na(Cancer_Status) & is.na(Cancer_Type)))
```

Observe that since the remaining $369  - 250 = 119$ observations still having `Cancer_Type == NA` also have that `Cancer_Status == 0`, meaning they do have cancer (thus setting `Cancer_Type` to $0$ is not a valid option) and hence we can either assign values to these missing values or once again, simply remove them.

-   Option 1: We calculate the estimate probability of each category (other than $0$) based on their frequency in the given data, and then assign `NA` values using our new probability measure.

```{r}
set.seed("14")

# calculate the estimated prob. of each category (except 0, and NAs)
cancer_status_category_prob <- brownfat_cleaned_data %>%
  filter(!is.na(Cancer_Type) & Cancer_Type != 0) %>%
  count(Cancer_Type) %>% 
  mutate(probability = n / sum(n))

# function to re-assign those w/ NA values
assign_cancer_type <- function(x) {
  if (is.na(x)) {
    sample(cancer_status_category_prob$Cancer_Type, 1, 
           prob = cancer_status_category_prob$probability)
  } else {
    x
  }
}


# re-assign those w/ Cancer_Type == NA
brownfat_cleaned_data_v1 <- brownfat_cleaned_data
brownfat_cleaned_data_v1$Cancer_Type <- sapply(brownfat_cleaned_data$Cancer_Type,
                                   assign_cancer_type)
```

Why may this method be problematic? =\> We've only taking into the account the frequencies of each of these types of cancer when in reality, a multitude of other factors affect the diagnosis of single patient. (We may consider to improve this later on by also creating a prediction model)

-   Option 2: To avoid potential hazardous assumptions, we may opt to exclude these `NA` values altogether:

```{r}
brownfat_cleaned_data_v2 <- brownfat_cleaned_data %>%
  filter(!is.na(Cancer_Type))
```

Lets check for the distribution of the age:
```{r}
ggplot(data = brownfat_cleaned_data_v2, aes(x = Age))+
  geom_histogram()
ggplot(data = brownfat_cleaned_data_v2, aes(x = Ext_Temp))+
  geom_histogram()
ggplot(data = brownfat_cleaned_data_v2, aes(x = `2D_Temp`))+
  geom_histogram()
ggplot(data = brownfat_cleaned_data_v2, aes(x = `3D_Temp`))+
  geom_histogram()
ggplot(data = brownfat_cleaned_data_v2, aes(x = `7D_Temp`))+
  geom_histogram()
ggplot(data = brownfat_cleaned_data_v2, aes(x = `1M_Temp`))+
  geom_histogram()
ggplot(data = brownfat_cleaned_data_v2, aes(x = Duration_Sunshine))+
  geom_histogram()
ggplot(data = brownfat_cleaned_data_v2, aes(x = Weigth))+
  geom_histogram()
ggplot(data = brownfat_cleaned_data_v2, aes(x = Size))+
  geom_histogram()
ggplot(data = brownfat_cleaned_data_v2, aes(x = BMI))+
  geom_histogram()
ggplot(data = brownfat_cleaned_data_v2, aes(x = Glycemy))+
  geom_histogram()
ggplot(data = brownfat_cleaned_data_v2, aes(x = LBW))+
  geom_histogram()
ggplot(data = brownfat_cleaned_data_v2, aes(x = Total_vol))+
  geom_histogram()
```

```{r}
mean_age <- mean(brownfat_cleaned_data_v2$Age)
mean_age
```
- The explanatory variable Age seems to follow approximately close to a normal distribution with a mean of 61.82. The sample looks skewed toward the elderly side of the adults.
```{r}
mean_ext_temp <- mean(brownfat_cleaned_data_v2$Ext_Temp)
mean_ext_temp
```
- The explanatory variable Ext_Temp seems to follow approximately but far from a normal distribution, with a mean of 6.

The correlation matrix of the model is given by: 
```{r}
brownfat_cleaned_data_v2$Month <- as.numeric(factor(brownfat_cleaned_data_v2$Month))
cor_matrix <- cor(brownfat_cleaned_data_v2)
pheatmap(cor_matrix, display_numbers = TRUE, cluster_rows = FALSE, cluster_cols = FALSE)

```
We notice that:
- All the temperature covariates are strongly, positively correlated to each other. 
- Weight is highly correlated with BMI and LBW.
- Sex, Size and LBW are strongly, positively  correlated to each other.
- The season is negatively, moderately correlated to the temperature and sunshine duration covariates.
- The sunshine is positively, strongly correlated to the temperature covariates.
- Cancer status and cancer type are positively, moderately correlated to each other.

Lets try centering the model:

```{r}
brownfat_cleaned_data_v2$Ext_Temp <- brownfat_cleaned_data_v2$Ext_Temp - mean(brownfat_cleaned_data_v2$Ext_Temp)

brownfat_cleaned_data_v2$`2D_Temp` <- brownfat_cleaned_data_v2$`2D_Temp` - mean(brownfat_cleaned_data_v2$`2D_Temp`) 

brownfat_cleaned_data_v2$`3D_Temp` <- brownfat_cleaned_data_v2$`3D_Temp` - mean(brownfat_cleaned_data_v2$`3D_Temp`) 

brownfat_cleaned_data_v2$`7D_Temp` <- brownfat_cleaned_data_v2$`7D_Temp` - mean(brownfat_cleaned_data_v2$`7D_Temp`)

brownfat_cleaned_data_v2$`1M_Temp` <- brownfat_cleaned_data_v2$`1M_Temp` - mean(brownfat_cleaned_data_v2$`1M_Temp`) 

brownfat_cleaned_data_v2$Season <- brownfat_cleaned_data_v2$Season - mean(brownfat_cleaned_data_v2$Season) 
brownfat_cleaned_data_v2$Duration_Sunshine<- brownfat_cleaned_data_v2$Duration_Sunshine - mean(brownfat_cleaned_data_v2$Duration_Sunshine) 

cor_matrix <- cor(brownfat_cleaned_data_v2)
pheatmap(cor_matrix, display_numbers = TRUE, cluster_rows = FALSE, cluster_cols = FALSE)

```